---
title: "Final Project (CI, HT, and LR)"
author: "Anthony Ramirez-Mata"
date: "2024-05-16"
output: html_notebook
---


# Find this code also on my GitHub "staticburnttoast"
# https://github.com/staticburnttoast/DATA105-Project


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# inject data
# data is listed and categorised into columns and rows
```{r}
test_data = read.csv("/Users/imnot/Downloads/test_results.csv")

head(test_data)
show(test_data)
```

# data details
# gives a `summary` of each column, such as minimums and maximums, quartiles, and mean medians and modes
# identifies the number of observations in a column and their type
```{r}
summary(test_data)
str(test_data)
```

# column and rows
# number of columns and rows, the names of columns
```{r}
ncol(test_data) # columns
nrow(test_data) # rows
colnames(test_data) # column titles
```

# column statistic of a specific column -> test_score
# lists the statistical information of a chosen column
```{r}
mean(test_data$test_score)
median(test_data$test_score)
sd(test_data$test_score)
var(test_data$test_score)
sum(test_data$test_score)
range = max(test_data$test_score) - min(test_data$test_score)
print(range)
quantile(test_data$test_score)
coefficientOfVariation = sd(test_data$test_score) / mean(test_data$test_score)
print(coefficientOfVariation)
z_scores = (test_data$test_score - mean(test_data$test_score)) / sd(test_data$test_score)
print(z_scores)
```

# correlation and covariance of two columns -> yrs_experience & salary
# lists the correlation and covariance between two chosen columns
```{r}
cor(test_data$yrs_experience, test_data$salary)
cov(test_data$yrs_experience, test_data$salary)
```

# correlation graph of data columns
# creates a graph based on the correlation of two columns
# the more blue and larger the dot, the greater the correlation
# the more red and smaller the dot, the lower the correlation
```{r}
library(corrplot)

M2 = cor(test_data[,c("yrs_experience", "test_score", "salary")])
corrplot(M2, method = "number")
corrplot(M2)
```

# histogram and skewness value
# creates a histogram and identifies the frequency of a chosen column
# prints the skewness of a column
```{r}
hist(test_data$test_score, xlab = "test score", ylab = "frequency", main = "test score histogram", col = "pink", labels = TRUE)

library(moments)
skewness(test_data$test_score)
```


# scatter plot
# creates a scatter plot of two chosen columns
```{r}
plot(test_data$test_score, test_data$salary, xlab = "test score", ylab = "salary", col = "blue")
```

# scatter plot with linear regression line
```{r}
plot(y = test_data$test_score, x = test_data$salary, xlab = "salary", ylab = "test_score", main = "the relationship between salary and test scores", pch = 16)
abline(lm(test_score ~ salary, data = test_data), col = "blue")
```

# 3D scatter plot
```{r}
library(scatterplot3d)

scatterplot3d(x = test_data$yrs_experience, y = test_data$test_score, z = test_data$salary,
              color = "blue",
              pch = 16,
              xlab = "years experience",
              ylab = "test scores",
              zlab = "salary",
              main = "3D scatterplot of the test data")
```

# box plot
```{r}
boxplot(test_data[,c("yrs_experience", "salary")])
```

# violin plot
```{r}
library(vioplot)
vioplot(test_data$yrs_experience, test_data$test_score, test_data$salary)
```

# density plot
```{r}
library(lattice)
densityplot(~test_data$test_score, main = "density plot", xlab = "salary", col = "blue")

plot(density(test_data$yrs_experience), col = "purple")
lines(density(test_data$salary), col = "pink")
```

# categorical column to data
```{r}
test_data$category = "high"
test_data$category[test_data$test_score < median(test_data$test_score)] = "low"
test_data$category[test_data$test_score == max(test_data$test_score)] = "max"
show(test_data)
```

# frequency table
```{r}
freq_tbl_rate_cat = table(test_data$category)
print(freq_tbl_rate_cat)
```

# barplot from frequency table
```{r}
barplot(freq_tbl_rate_cat, xlab = "categories of test data", ylab = "frequency", main = "test data categories")
```

# pie chart
```{r}
pie(freq_tbl_rate_cat)
pie(freq_tbl_rate_cat, paste0(freq_tbl_rate_cat, "%"), col = "violet")
```

# confidence intervals
```{r}
model = lm(test_score~1, test_data)
confint(model, data=test_data)
```

# hypothesis testing
```{r}
t.test(test_data$test_score, mu = 700, conf.level = .95)
```

# hypothesis testing pt.2
```{r}
t.test(test_data$test_score, test_data$salary, conf.level = .95)
```

# linear regression model
```{r}
lr_model = lm(yrs_experience ~ salary, data = test_data)
summary(lr_model)
```

# predictions on model
# with a salary of 60, we can see approximately 16 years of experience, while a salary of 100 equals 31 years of experience. (wow...)
```{r}
data_prediction = data.frame(salary = c(60, 100))

predict(lr_model, data_prediction)
```

# graphing linear regression
```{r}
plot(test_data$salary, test_data$yrs_experience, 
     main = "Linear Regression", 
     xlab = "Salary", ylab = "Years of Experience",
     pch = 19, col = "orange")

abline(lr_model, col = "cyan")
```
